% =========================================================================
% CAVE-HAR — 3DV 2027 draft (target deadline: Aug 28, 2026 AoE)
% -------------------------------------------------------------------------
% NOTE: 3DV uses the CVF-style two-column format. Before submission,
% replace this preamble with the official 3DV 2027 author kit (expected on
% https://3dvconf.github.io/2027/call-for-papers/) and delete the \todo
% macro output by switching \todofalse. All open gaps are marked \todo{...}
% and indexed in ../GAPS.md with owner + producing phase.
% =========================================================================
\documentclass[10pt,twocolumn,letterpaper]{article}

\usepackage[margin=1in]{geometry}   % replaced by official style file
\usepackage{times}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{amsmath}
\usepackage{pifont}   % \ding crosses in the dataset comparison table
\usepackage{xcolor}
% Draft previews contain Chinese \todo notes: build with XeLaTeX, where the
% todo macro switches to a CJK font (fontspec only, no xeCJK dependency).
% The final submission (todos removed via \todosfalse) compiles with the
% official pdflatex-based 3DV kit as-is.
\usepackage{iftex}
\ifXeTeX
  \usepackage{fontspec}
  \newfontfamily\todocjkfont{Noto Serif CJK SC}
\else
  \newcommand{\todocjkfont}{}
\fi
\usepackage[pagebackref,breaklinks,colorlinks,citecolor=blue]{hyperref}

% ---- draft machinery ----------------------------------------------------
\newif\iftodos \todostrue   % \todosfalse hides all TODO markers
\newcommand{\todo}[1]{\iftodos\textcolor{red}{\textbf{[TODO: {\todocjkfont #1}]}}\fi}
\newcommand{\DATASET}{CAVE-HAR\todo{数据集正式命名待定 (G1)}}
\newcommand{\numclips}{\todo{清洗后片段总数 (G8)}}

\begin{document}

\title{\DATASET: A Multi-View Benchmark for Human Action Recognition\\
in Immersive CAVE Environments}

\author{Anonymous 3DV 2027 submission\\Paper ID \todo{****}}
\date{}
\maketitle

% =========================================================================
\begin{abstract}
Human action recognition (HAR) inside large-scale immersive systems such as
CAVEs (Cave Automatic Virtual Environments) promises marker-less
interaction, automatic procedure assessment, and behavioral analysis for
safety-critical training. Yet every widely used HAR benchmark implicitly
assumes conditions that a CAVE violates: frontal tripod-height viewpoints,
stable photometric backgrounds, and offline evaluation. We present
\DATASET{}, the first HAR benchmark captured inside an operational
projection-based CAVE. It provides synchronized multi-view RGB video from
seven cameras---including the top-mounted, steeply pitched views forced by
CAVE geometry---together with optical-motion-capture-grade 3D skeleton
ground truth (SMPL layout, 22 joints) for 15 actors performing 7 general
and 11 VR-interaction actions across 6 projected scenes. Because the
projected VR content changes the entire visual background under controlled
conditions, \DATASET{} additionally enables a background-shift evaluation
that is impossible to stage in real-world scenes. We define four evaluation
protocols (cross-subject, cross-view, cross-scene, and cross-domain from
NTU~RGB+D~120), plus a latency-aware protocol that reports computation and
observation latency for streaming recognition. Benchmarking
\todo{N (G13)} representative skeleton, video, and multimodal models
reveals \todo{核心发现一句话, 需实验后填 (G13)}. The dataset, evaluation
code, and correction toolchain will be publicly released.
\end{abstract}

% =========================================================================
\section{Introduction}
Projection-based immersive systems such as the CAVE~\cite{cruzneira1993cave}
remain the platform of choice when virtual reality must be combined with
free full-body movement, unobstructed vision of one's own body, and
multi-user co-presence---requirements typical of industrial safety
training, rehabilitation, and simulation. Human action recognition inside
such environments would enable automatic evaluation of trainee procedures,
marker-less interaction, and quantitative behavioral analysis.

Modern HAR models are, however, trained and evaluated almost exclusively on
datasets that encode three assumptions. First, \emph{fixed frontal
geometry}: cameras in NTU~RGB+D~\cite{shahroudy2016ntu,liu2020ntu120} and
its successors observe subjects from tripod height at moderate distances,
whereas a CAVE forces cameras to its ceiling rim, producing steep pitch
angles and systematic self-occlusion of hands and head during interaction.
Second, \emph{appearance stability}: laboratory backgrounds are static,
whereas a CAVE's walls and floor \emph{are} the display---the background
changes radically with the projected scene and its light spills onto the
actor. Third, \emph{offline evaluation}: accuracy is reported on trimmed
clips without latency constraints, whereas interactive VR imposes tight
motion-to-response budgets. The gap between these assumptions and CAVE
reality is unquantified because no suitable benchmark exists.

We close this gap with \DATASET{}, captured in a
\todo{CAVE 尺寸/面数确认: 幻灯片同时出现 ``5 faces active'' 与 ``6-sided''
(G2)} projection CAVE instrumented with 12 OptiTrack Prime$^{\mathrm{x}}$13
motion-capture cameras and 7 RGB cameras. Our contributions are:
\begin{itemize}
  \item \textbf{Dataset.} \numclips{} clips of 15 actors (9M/6F), 6
        projected scenes (3 indoor, 3 outdoor), 7 general + 11
        VR-interaction actions, with synchronized 7-view RGB and
        mocap-grade SMPL-22 skeletons---to our knowledge the first HAR
        corpus recorded inside an operational CAVE, and the first with
        top-mounted viewpoints as a primary condition.
  \item \textbf{Human-verified ground truth.} A correction pipeline
        combining per-clip quality scoring, two-view triangulated editing,
        and bone-length-preserving inverse-kinematics tools; all released
        test-split skeletons are manually verified
        (\todo{修正帧占比 X\% (G9)}, inter-annotator agreement
        \todo{Y mm (G10)}).
  \item \textbf{Protocols.} Cross-subject, cross-view, cross-scene, and
        NTU$\rightarrow$\DATASET{} cross-domain splits, plus a
        latency-aware streaming protocol reporting computation and
        observation latency.
  \item \textbf{Baselines.} \todo{N (G13)} models spanning skeleton GCNs
        and transformers, video networks, and multimodal fusion, with an
        analysis of where and why they fail under CAVE conditions.
\end{itemize}

% =========================================================================
\section{Related Work}
\paragraph{HAR datasets.}
Existing corpora split into two families with an inherent
scale--fidelity trade-off. The \emph{scale} family---NTU~RGB+D
\cite{shahroudy2016ntu}, NTU~RGB+D~120~\cite{liu2020ntu120},
ANUBIS~\cite{yu2022anubis}, Kinetics~\cite{kay2017kinetics}---recruits
large cohorts but estimates skeletons from consumer depth sensors or 2D
pose, with joint errors on the order of centimeters and frequent tracking
failure under occlusion. The \emph{fidelity} family anchors joints with
optical motion capture and has always traded cohort size for precision:
Human3.6M~\cite{ionescu2014h36m} records 11 actors with 4 cameras,
TotalCapture~\cite{trumble2017totalcapture} 5 actors with 8 cameras.
\DATASET{} belongs to the fidelity family---240\,Hz optical mocap refined
by MoSh++ and manual verification---while providing the largest actor
cohort in that family, the most action classes, and conditions none of
them cover: top-mounted viewpoints, prop-based VR interaction, and
controlled background shift (Table~\ref{tab:datasets}).

\paragraph{User behavior corpora in VR.}
LocoVR~\cite{takeyama2024locovr} records multi-user locomotion trajectories
in HMD-based VR; HUMOTO~\cite{lu2025humoto} provides mocap human--object
interactions without multi-view RGB. HMD-based corpora observe users
wearing headsets that occlude the face and alter motion; CAVE users wear
only lightweight shutter glasses, keeping appearance close to natural while
introducing the projection-specific challenges we target. RoCoG-v2
\cite{reddy2023rocog} studies synthetic-to-real domain shift for gestures,
methodologically related to our cross-domain protocol.

\paragraph{Models.}
Skeleton-based recognition evolved from graph convolutions
(ST-GCN~\cite{yan2018stgcn}, CTR-GCN~\cite{chen2021ctrgcn}) to
attention and prototype-based designs (SkateFormer~\cite{do2024skateformer},
ProtoGCN~\cite{liu2025protogcn}), state-space models
(SkelMamba~\cite{martinel2024skelmamba}), and pretrained motion encoders
(MotionBERT~\cite{zhu2023motionbert}). Video models progressed from
two-pathway CNNs (SlowFast~\cite{feichtenhofer2019slowfast}) to
transformers (Video Swin~\cite{liu2022videoswin},
VideoMAE~v2~\cite{wang2023videomae2}) and state-space architectures
(VideoMamba~\cite{li2024videomamba}); language--vision pretraining enables
zero-shot transfer (ActionCLIP~\cite{wang2021actionclip}); and
PoseConv3D~\cite{duan2022poseconv3d} fuses pose heatmaps with RGB. We
benchmark representatives of each family.

\paragraph{Online and latency-aware evaluation.}
Online action detection~\cite{degeest2016oad} and streaming perception
argue that offline accuracy overstates deployable performance; recent work
separates \emph{computation} latency from \emph{observation} latency
\cite{eyzaguirre2024squid}. VR interaction makes these constraints
first-class, motivating our protocol.

% =========================================================================
\section{The \DATASET{} Dataset}

\subsection{Capture environment}
Recording took place in the \todo{机构匿名化处理后的 CAVE 描述 + 尺寸
(G2)} CAVE. Twelve OptiTrack Prime$^{\mathrm{x}}$13 cameras capture 3D
motion at 240\,Hz; seven RGB cameras (\todo{型号/分辨率/帧率 (G3)},
nominally 30\,fps) cover the volume in a $3{\times}2$ grid on the left,
center, and right rim at high and low mounting positions, plus one
diagonal rear view. Fig.~\ref{fig:rig}\todo{相机布置图 (G4)} shows the
layout; the ``high'' row is top-mounted at pitch angles of
\todo{俯仰角度数 (G3)}, the geometry that distinguishes CAVE deployments
from tripod-based datasets.

Two collection rounds were performed. Round~1 (portal-configuration LED
CAVE, \todo{7 或 8 (幻灯片不一致, G5)} mocap + 2 RGB cameras) served as a
pilot; Round~2, described above, is the release configuration. Actors wore
everyday clothing with 37 retro-reflective markers; Round~2 used low-profile
marker stickers to minimize visual interference with the RGB streams.

\subsection{Actors, scenes, and actions}
Fifteen actors (9 male, 6 female; \todo{年龄范围/身高范围, 需 roster (G6)})
each performed in six projected scenes (three indoor, three outdoor),
approximately 20 effective minutes per scene and roughly 41{,}400 mocap
frames per actor. The action vocabulary (Table~\ref{tab:actions}) contains
seven general actions shared with existing benchmarks---enabling direct
cross-domain comparison---and eleven VR-interaction actions elicited by the
projected applications, from controller locomotion to catching virtual
fish with a physical net prop.

\begin{table}[t]\centering\small
\caption{Action vocabulary: 7 general + 11 VR-interaction classes.}
\label{tab:actions}
\begin{tabular}{@{}ll@{}}
\toprule
\textbf{General} & \textbf{VR-interaction} \\
\midrule
walking      & move via controller \\
running      & catch fish with net \\
jumping      & grab box \\
bending down & measure length \\
standing     & wave \\
squatting    & throw \\
raising hand & cut \\
             & shoot \\
             & pick up item \\
             & go up stairs \\
             & dodge car \\
\bottomrule
\end{tabular}
\end{table}

\begin{table*}[t]\centering\small
\caption{Comparison with representative HAR and motion-capture datasets.
Estimated-skeleton corpora (top) scale in subjects but provide noisy
joints; mocap-grade corpora (bottom) trade cohort size for precision.
\DATASET{} offers the largest cohort and action vocabulary in the
mocap-grade family, with capture conditions unavailable elsewhere.
\todo{逐格终核数字 (G7)}}
\label{tab:datasets}
\begin{tabular}{@{}lcccccccc@{}}
\toprule
Dataset & Subjects & Classes & Skeleton GT (rate) & Calib.\ RGB views &
Top-mounted & Props/VR interact. & Ctrl.\ bg shift & Verified GT \\
\midrule
\multicolumn{9}{@{}l}{\emph{Estimated skeletons (scale family)}} \\
NTU RGB+D 120~\cite{liu2020ntu120} & 106 & 120 & Kinect v2 (30\,Hz) & 3 viewpoints & \ding{55} & \ding{55} & \ding{55} & \ding{55} \\
ANUBIS~\cite{yu2022anubis} & 80 & 102 & Azure Kinect (30\,Hz) & \ding{55} & \ding{55} & \ding{55} & \ding{55} & \ding{55} \\
\midrule
\multicolumn{9}{@{}l}{\emph{Mocap-grade skeletons (fidelity family)}} \\
Human3.6M~\cite{ionescu2014h36m} & 11 & 17 scen. & optical mocap (50\,Hz) & 4 & \ding{55} & \ding{55} & \ding{55} & \ding{55} \\
TotalCapture~\cite{trumble2017totalcapture} & 5 & 5 acts & Vicon+IMU (60\,Hz) & 8 & \ding{55} & \ding{55} & \ding{55} & \ding{55} \\
\DATASET{} (ours) & 15 & 18 & optical mocap (240\,Hz) + MoSh++ & 7 & \checkmark & \checkmark & \checkmark & \checkmark \\
\bottomrule
\end{tabular}
\end{table*}

\subsection{Calibration and synchronization}
RGB intrinsics and extrinsics follow Zhang's method~\cite{zhang1999calib}
with per-session refinement; extrinsic drift is monitored by
re-projecting triangulated 2D poses and gating on median residual
(\todo{标定精度: 平均重投影误差 px (G11)}). Lacking hardware genlock,
video--skeleton alignment uses a per-clip timeline established from the
mocap clock, refined per camera with sub-frame offsets and verified by
overlay inspection; final alignment accuracy is \todo{同步精度 ±帧 (G11)}.

\subsection{Ground-truth construction}
Raw 240\,Hz markers are solved to SMPL~\cite{loper2015smpl} bodies with
MoSh++~\cite{mahmood2019amass}; we release the 22-joint SMPL subset
(hand joints excluded owing to sparse hand markers) in both world
coordinates and pelvis-centered normalized form. Sparse marker coverage
and interaction props cause characteristic failures---frozen or drifting
limb segments---which we repair with a purpose-built toolchain: (i) every
clip receives an automatic quality score combining bone-length deviation,
robust velocity-spike detection, frozen-joint-while-moving runs, and
missing-marker fraction, which prioritizes review worst-first;
(ii) annotators correct joints by dragging in two calibrated views
simultaneously (exact triangulated depth) with keyframe--spline
propagation; and (iii) a bone-length-preserving two-bone IK mode solves
whole-limb placements from a single wrist/ankle drag, guaranteeing
physically plausible corrections that never fold a straight limb. All
test-split clips are fully verified; training-split clips are corrected
worst-first by quality score with per-frame edit masks released
(\todo{银标准修正比例 (G9)}). Corrections modified \todo{X\% (G9)} of
test-split frames; dual-annotation of a 5\% sample shows mean joint
disagreement of \todo{Y mm (G10)}. Fig.~\todo{标注工具界面图: 双视角+极线
+IK黄圈, 一图讲清方法论 (G22)} illustrates the tool; the full correction
SOP, marker layout, and capture protocol appear in the supplementary
material, and the toolchain itself is released so that others can extend
the corpus at the same quality bar.

\subsection{Quantifying the CAVE gap}
\label{sec:gap}
To turn the three violated assumptions of \S1 into measurements, we
characterize the corpus along three axes; all quantities are computed
directly from the released calibration and ground truth.
\paragraph{Viewpoint geometry.}
From the camera extrinsics we report the distribution of camera pitch
angles relative to the subject's pelvis across all frames: the top-mounted
row observes actors at \todo{中位俯仰角/范围, 由外参+骨架直接算 (G19)},
versus the near-horizontal geometry of tripod-based corpora.
\paragraph{Interaction-driven self-occlusion.}
Projecting the mocap ground truth into each view and testing body
self-occlusion \todo{遮挡判定方法一句话 (G20)}, wrists are occluded in
\todo{X\% (G20)} of interaction-action frames in top-mounted views versus
\todo{Y\% (G20)} in the low-rim views---occlusion statistics that
ground-truth-free datasets cannot even measure.
\paragraph{Projected-background variation.}
Because the CAVE walls are displays, scene changes shift the entire
background and spill colored light onto the actor. We quantify this as the
inter-scene distance of background color statistics and the intra-actor
appearance drift across scenes (\todo{LAB 直方图距离等具体度量与数字
(G21)}), an order of magnitude \todo{核实后措辞 (G21)} beyond the
illumination variation of laboratory corpora.

\subsection{Statistics}
\todo{最终统计表: 每类样本数/时长分布/marker 丢失率/QC 分布直方图,
清洗后生成 (G8)}

% =========================================================================
\section{Benchmark Protocols}
\paragraph{Cross-subject (X-Sub, primary).}
Three fixed 10-train/5-test actor partitions (seeded, gender-balanced
\todo{性别核对后确认 (G12)}); we report mean$\pm$std over the three
partitions to control for small-cohort variance.
\paragraph{Cross-view (X-View).}
Train on low-rim views, test on top-mounted and diagonal-rear views,
isolating the viewpoint gap.
\paragraph{Cross-scene (X-Scene).}
Train on the three indoor projections, test on the three outdoor ones:
identical actions, radically different projected backgrounds---a
controlled background-shift experiment unavailable in real-world corpora.
\paragraph{Cross-domain.}
Models pretrained on NTU RGB+D 120 are evaluated zero-shot and after
fine-tuning, quantifying the lab-to-CAVE domain gap that motivates this
work.
\paragraph{Metrics.}
Top-1/Top-5 accuracy and mean per-class accuracy (the 18-class vocabulary
is imbalanced across interaction types); confusion analyses separate
general from interaction classes.
\paragraph{Latency-aware protocol.}
For streaming deployment we report (i)~\emph{computation latency}:
single-clip inference time on one \todo{评测 GPU 型号 (G14)}, and
(ii)~\emph{observation latency}: the minimum observed-frame budget at
which a model reaches 90\% of its full-clip accuracy
\cite{degeest2016oad,eyzaguirre2024squid}, yielding accuracy--latency
curves (Fig.~\todo{latency 图 (G15)}).

% =========================================================================
\section{Experiments}
\todo{整节数字待 P3 实验产出 (G13). 表格骨架已按分工排好.}

\begin{table}[t]\centering\small
\caption{Baselines on \DATASET{} (Top-1 / mean per-class, \%).
\todo{全部数字 (G13)}}
\label{tab:baselines}
\begin{tabular}{@{}llccc@{}}
\toprule
Model & Modality & X-Sub & X-View & X-Scene \\
\midrule
ST-GCN~\cite{yan2018stgcn}             & Skel. & \todo{} & \todo{} & \todo{} \\
CTR-GCN~\cite{chen2021ctrgcn}          & Skel. & \todo{} & \todo{} & \todo{} \\
SkateFormer~\cite{do2024skateformer}   & Skel. & \todo{} & \todo{} & \todo{} \\
SkelMamba~\cite{martinel2024skelmamba} & Skel. & \todo{} & \todo{} & \todo{} \\
MotionBERT~\cite{zhu2023motionbert}    & Skel. & \todo{} & \todo{} & \todo{} \\
SlowFast~\cite{feichtenhofer2019slowfast} & RGB & \todo{} & \todo{} & \todo{} \\
Video Swin~\cite{liu2022videoswin}     & RGB   & \todo{} & \todo{} & \todo{} \\
VideoMAE v2~\cite{wang2023videomae2}   & RGB   & \todo{} & \todo{} & \todo{} \\
VideoMamba~\cite{li2024videomamba}     & RGB   & \todo{} & \todo{} & \todo{} \\
ActionCLIP~\cite{wang2021actionclip}   & RGB+T & \todo{} & \todo{} & \todo{} \\
PoseConv3D~\cite{duan2022poseconv3d}   & Skel.+RGB & \todo{} & \todo{} & \todo{} \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Cross-domain gap.} \todo{NTU 预训练 zero-shot/finetune 表 (G13)}
\paragraph{Viewpoint analysis.} Per-camera accuracy correlated with the
pitch-angle and occlusion statistics of \S\ref{sec:gap}.
\todo{逐相机准确率 + 遮挡消融 (G13)}
\paragraph{Per-class analysis.} Confusion between general and
VR-interaction classes; which interaction actions are separable from
skeleton alone versus requiring RGB (prop visibility).
\todo{混淆矩阵与讨论 (G13)}
\paragraph{Does ground-truth quality matter?}
Training the strongest skeleton baseline on raw MoSh++ output versus
verified skeletons isolates the value of manual correction
(\todo{原始 vs 修正 GT 消融, 训练两次即可, 低成本高说服力 (G23)})---an
ablation that also serves practitioners deciding whether correction effort
is worthwhile.
\paragraph{Zero-shot recognition.} \todo{ActionCLIP 零样本结果 (G13)}
\paragraph{Latency.} \todo{accuracy-latency 曲线与讨论 (G13, G15)}
\paragraph{Qualitative results.} \todo{成功/失败案例可视化 (G16)}

% =========================================================================
\section{Limitations and Ethics}
\DATASET{} is captured in a single CAVE installation with 15 actors; the
cohort size follows from mocap-suit logistics and is mitigated by the
multi-partition X-Sub protocol, but broader demographic coverage is future
work. Hand joints are excluded from the official set because sparse hand
markers make them unreliable. Training-split skeletons are corrected
worst-first rather than exhaustively; per-frame edit masks make the
verification status of every joint transparent. All participants gave
written informed consent covering public release of RGB video and derived
data under an institution-approved protocol
(\todo{伦理批号, 投稿系统需要 (G17)}); raw uncorrected captures are not
distributed. \todo{面部是否模糊处理的最终决定 (G17)}

\section{Conclusion}
\DATASET{} turns an operational CAVE into a benchmark, pairing
mocap-grade skeletons with the viewpoints, backgrounds, and latency
constraints that immersive deployment actually imposes. Our baselines show
\todo{一句话结论 (G13)}, and the released toolchain lowers the cost of
extending the corpus. We hope \DATASET{} seeds a line of HAR research that
treats immersive environments as a first-class setting rather than an
afterthought.

{\small
\bibliographystyle{ieee_fullname}
\bibliography{refs}
}

\end{document}
